Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-28 VMCAI-revised-version/final.tex
effectivity

\begin{center}
 \resizebox{\textwidth}{!}{
\begin{tabular}{ | l | c | c | c | r |}
\hline   \ 
		& {\bf WBTS model (with $\Safe = \mathop{\downarrow} \Safe$)} ~ 
		 & ~ {\bf Decidability}  ~ 
 \\ \hline
   \multirow{3}{*}{ {\bf RP }} 
   		 & Downard-WSTS ($\pred^*$-basis-effective) Thm~\ref{up-down}
   		 	& \multirow{2}{*}{Decidable}  \\

@@

   \multirow{2}{*}{{\bf sRP ~ ~ }} 
   		 & WSTS (downward + upward) Thm~9 & Decidable\\ \cline{2-3}
~	& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down}  & Undecidable  \\ \hline
   \multirow{2}{*}{{\bf BsRP }} 
   		 & WSTS (downward + upward) Cor~\ref{downward brp} & Decidable \\ \cline{2-3}
~	&  ? & Undecidable   \\ \hline
   \multirow{2}{*}{{\bf ksRP }} 
   		 & WSTS (downward + upward) Cor~\ref{downward brp} & Decidable\\ \cline{2-3}
~	 & ? & Undecidable    \\ \hline

@@

Another type of WSTS exists that enjoys downward compatibility.
%
%		Finkel and Schnoebelen, 
%
\begin{definition}[Definition 5.1, \cite{DBLP:journals/tcs/FinkelS01}]
A {\em downward (compatible) well-structured transition system} (downward-WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ such that   
the transition relation $ \rightarrow$ is downward compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_2 \leq t_1$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

@@

%
\begin{theorem}\label{up-down}
{\sc Resilience}  is decidable for $\pred^*$-basis-effective downward-WSTS with 
$\Safe=\mathop{\downarrow} \Safe$.
\end{theorem}

\begin{proof}
Let $\mathscr{S}$ be a $\pred^*$-basis-effective downward-WSTS with 
$\Safe=\mathop{\downarrow} \Safe$.
By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed; moreover, $\pred^*(\Safe)$ is computable because $\mathscr{S}$ is $\pred^*$-basis-effective.
The resilience problem can be reformulated as $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
%\alain{mais comment je calcule $\pred^*(\Safe)$ ?}
Since $\leq$ is ideally-effective, we can compute complementaries and intersections of upward and downward-closed
 subsets.
Hence we can compute $S \setminus \pred^*(\Safe)$ and then the intersection of
$\Bad$
and
$S \setminus \pred^*(\Safe)$,
which are both upward-closed. We may decide whether this intersection is empty, hence
the resilience problem is decidable. \qed
\end{proof}

In the case
of a $\pred^*$-basis-effective WSTS, not necesarly downward compatible,
the above construction
can provide a proof
of non-resilience
i.e. when
$\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) \neq \emptyset$
then
$\Bad \not\subseteq \mathop{\downarrow}\pred^*(\Safe)$
and hence
$\Bad \not\subseteq \pred^*(\Safe)$.
When $\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) = \emptyset$
however
it is not enough to conclude.\\

@@

%\alain{énoncer pour les reset-VASS c'est plus fort}
\begin{proposition}\label{srp up down}
{\sc State-resilience},
{\sc Bounded-state-resilience} and
{\sc $k$-state-resilience}
are undecidable for reset-VASS, hence for strongly compatible,
$\post$-effective, $\pred$-basis-effective WSTS
when
$\Safe=\mathop{\uparrow} \Safe$.
\end{proposition}

@@

In the case of a WSTS with \emph{downward} compatibility, not necessarily strong,
then $\Safe$ downward-closed implies $\pred^*(\Safe)$ downward-closed and
Lemma~\ref{Lemma intersection 2} can be used to show that
if $\Safe = \mathop{\downarrow} \Safe$,
then
$\post^*(s)  
\subseteq \pred^*(\Safe)$  iff $ (\mathop{\downarrow}  \post^*(s)) 
\subseteq \pred^*(\Safe)$.

\begin{theorem}\label{downward srp}
{\sc State-resilience} is decidable for $\post$-effective, 
%$\pred$-basis-effective, 
$\post$-ideal-effective and $\pred^*$-basis-effective WSTS with downward and upward compatibilities
and $\Safe = \mathop{\downarrow} \Safe$.
\end{theorem}

@@

\begin{corollary}\label{downward brp}
{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility when
$\Safe = \mathop{\downarrow} \Safe$
and there exists an algorithm accepting any ideal $I$ and returning a finite basis of
 $\mathop{\downarrow} \pred(I)$.
\end{corollary}

>>>>

\begin{center}
 \resizebox{\textwidth}{!}{
\begin{tabular}{ | l | c | c | c | r |}
\hline   \ 
		& {\bf WBTS model (with $\Safe = \mathop{\downarrow} \Safe$)} ~ 
		 & ~ {\bf Decidability}  ~ 
 \\ \hline
   \multirow{3}{*}{ {\bf RP }} 
   		 & Downard-WBTS ($\pred^*$-basis-effective) Thm~\ref{up-down}
   		 	& \multirow{2}{*}{Decidable}  \\

@@

   \multirow{2}{*}{{\bf sRP ~ ~ }} 
   		 & WBTS (downward + upward) Thm~9 & Decidable\\ \cline{2-3}
~	& ~ Reset-VASS, WSTS ($\pred$-basis-effective) Prop~\ref{srp up down}  & Undecidable  \\ \hline
   \multirow{2}{*}{{\bf BsRP }} 
   		 & WBTS (downward + upward) Cor~\ref{downward brp} & Decidable \\ \cline{2-3}
~	&  ? & Undecidable   \\ \hline
   \multirow{2}{*}{{\bf ksRP }} 
   		 & WBTS (downward + upward) Cor~\ref{downward brp} & Decidable\\ \cline{2-3}
~	 & ? & Undecidable    \\ \hline

@@

Another type of WSTS exists that enjoys downward compatibility.
%
%		Finkel and Schnoebelen, 
%
\begin{definition}
A {\em downward (compatible) WBTS (short, downward-WBTS)} $\mathscr{S}=(S, \rightarrow, \leq)$
is an ordered transition system $(S, \rightarrow, \leq)$ such that $\leq$ is without infinite antichains and
the transition relation $ \rightarrow$ is downward compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_2 \leq t_1$ and $s_2 \rightarrow^{*} t_2$.

A  {\em downward (compatible) WSTS (short, downward-WSTS)} \cite[Definition 5.1]{DBLP:journals/tcs/FinkelS01} is a downward-WBTS where $\leq$ is a wqo.
\end{definition}

@@

%
\begin{theorem}\label{up-down}
{\sc Resilience}  is decidable for $\pred^*$-basis-effective downward-WBTS with 
$\Safe=\mathop{\downarrow} \Safe$.
\end{theorem}

\begin{proof}
Let $\mathscr{S}$ be a $\pred^*$-basis-effective downward-WBTS with 
$\Safe=\mathop{\downarrow} \Safe$.
By Lemma \ref{downward compatible}, $\pred^*(\Safe)$ is downward-closed; moreover, $\pred^*(\Safe)$ is computable because $\mathscr{S}$ is $\pred^*$-basis-effective.
The resilience problem can be reformulated as $\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$.
%\alain{mais comment je calcule $\pred^*(\Safe)$ ?}
Since $\leq$ is ideally-effective, we can compute complementaries and intersections of upward and downward-closed
 subsets.
Hence we can compute $S \setminus \pred^*(\Safe)$ and then the intersection of
$\Bad$
and
$S \setminus \pred^*(\Safe)$,
which are both upward-closed. We may decide whether this intersection is empty, hence
the resilience problem is decidable. \qed
\end{proof}

In the case
of a $\pred^*$-basis-effective WBTS, not necesarly downward compatible,
the above construction
can provide a proof
of non-resilience
i.e. when
$\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) \neq \emptyset$
then
$\Bad \not\subseteq \mathop{\downarrow}\pred^*(\Safe)$
and hence
$\Bad \not\subseteq \pred^*(\Safe)$.
When $\Bad \cap (S \setminus \mathop{\downarrow}\pred^*(\Safe)) = \emptyset$
however
it is not enough to conclude.\\

@@

%\alain{énoncer pour les reset-VASS c'est plus fort}
\begin{proposition}\label{srp up down}
{\sc State-resilience},
{\sc Bounded-state-resilience} and
{\sc $k$-state-resilience}
are undecidable for reset-VASS, hence for strongly compatible,
$\post$-effective, $\pred$-basis-effective $\omega^2$-WSTS
when
$\Safe=\mathop{\uparrow} \Safe$.
\end{proposition}

@@

In the case of a WBTS with \emph{downward} compatibility, not necessarily strong,
then $\Safe$ downward-closed implies $\pred^*(\Safe)$ downward-closed and
Lemma~\ref{Lemma intersection 2} can be used to show that
if $\Safe = \mathop{\downarrow} \Safe$,
then
$\post^*(s)  
\subseteq \pred^*(\Safe)$  iff $ (\mathop{\downarrow}  \post^*(s)) 
\subseteq \pred^*(\Safe)$.

\begin{theorem}\label{downward srp}
{\sc State-resilience} is decidable for $\post$-effective, 
%$\pred$-basis-effective, 
$\post$-ideal-effective and $\pred^*$-basis-effective WBTS with downward and upward compatibilities
and $\Safe = \mathop{\downarrow} \Safe$.
\end{theorem}

@@

\begin{corollary}\label{downward brp}
{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\post$-effective, $\post$-ideal-effective WBTS, strong downward compatibility and upward compatibility when
$\Safe = \mathop{\downarrow} \Safe$
and there exists an algorithm accepting any ideal $I$ and returning a finite basis of
 $\mathop{\downarrow} \pred(I)$.
\end{corollary}
